Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-18 paper.tex
changed title
edit to intro

%% Title information
\title{Blame and Coercion: Together Again for the Second Time}

@@

Around 2006, several groups of researchers proposed ways to integrate
dynamic typing and static typing, notably gradual typing
\citep{siek2006gradual}, hybrid typing \citep{flanagan2006hybrid},
migratory typing \citep{Tobin-Hochstadt:2006fk}, and multi-language
iteroperability \citep{Gray:2005ij,Matthews:2007zr}. Researchers
usually define the semantics of gradually typed languages by
translation to an intermediate language with casts, such as the blame
calculus \citep{wadler2009well} and other cast calculi
\citep{siek2009exploring}. Unfortunately, straightforward
implementations of casts on higher-order values (functions, objects,
etc.) impose significant runtime overheads that can change the
asymptotic time and space complexity of a program
\citep{herman2010space}. There are several known space-efficient cast
representations, with various strengths and weaknesses. This paper
presents a new cast representation, named \emph{hypercoercions}, that
is good for both mechanized metatheory and good for implementation.
The composition operator for hypercoercions is structurally recursive
and the hypercoercions are suggestive of a bit-level representation
that minimizes the need for pointers and fits all first-order casts
into 64 bits.  Of course, valid cast representation must not change
the result of programs. This paper presents first steps toward a
framework for proving the equivalence of cast calculi.

@@

There are two best studied ways to represent cast space-efficiently: coercions 
\citep{herman2010space}\citep{siek2009exploring}\citep{siek2012interpretations} 
\citep{siek2015blame} and threesomes 
\citep{siek2010threesomes}\citep{garcia2013calculating}. They have different 
strength and 
weakness. \citet{garcia2013calculating} and \citet{siek2015blame} claim that 
coercion is easy to understand, but hard to implement, while threesome is easy 
to implement, but hard to understand. \citet{garcia2013calculating} in addition 
claims that coercion is more difficult to reason about formally because 
its compose function is not structurally recurisive. For instance, the 
developers of Grift \citep{kuhlenschmidt2018efficient} report that it is tricky 
to convince their proof assistants that the compose function terminates.

>>>>

%% Title information
\title{Hypercoercions and a Framework for Equivalence of Cast Calculi}

@@

Around 2006, several groups of researchers proposed ways to integrate
dynamic typing and static typing, notably gradual typing
\citep{siek2006gradual}, hybrid typing \citep{flanagan2006hybrid},
migratory typing \citep{Tobin-Hochstadt:2006fk}, and multi-language
iteroperability \citep{Gray:2005ij,Matthews:2007zr}. Researchers
usually define the semantics of gradually typed languages by
translation to an intermediate language with casts, such as the blame
calculus \citep{wadler2009well} and other cast calculi
\citep{siek2009exploring}. Unfortunately, straightforward
implementations of casts on higher-order values (functions, objects,
etc.) impose significant runtime overheads that can change the
asymptotic time and space complexity of a program
\citep{herman2010space}. There are several known space-efficient cast
representations, with various strengths and weaknesses
\citep{siek2015blame,siek2010threesomes,garcia2013calculating,kuhlenschmidt2018efficient,siek2012interpretations,garcia2014deriving}. This
paper presents a new cast representation, named \emph{hypercoercions},
that is good for both mechanized metatheory and good for
implementation. Recall that it is the composition operator in these
systems that enables space-efficiency by compressing casts.  The
composition operator for hypercoercions is structurally recursive and
hypercoercions are suggestive of a bit-level representation that
minimizes the need for pointers and fits all first-order casts into 64
bits.  Of course, an alternative cast representation must not change
the result of programs. This paper presents steps toward a framework
for proving the equivalence of cast calculi and, in particular, proves
that an abstract machine using hypercoercions is equivalent to an
abstract machine using standard casts.

@@

The current state of the art regarding cast representations includes
\begin{itemize}
\item threesomes \citep{siek2010threesomes},
\item supercoercions \citep{garcia2014deriving}, and
\item coercions in normal form
  \citep{siek2012interpretations,siek2015blame}.
\end{itemize}
Threesomes and supercoercions are good for mechanized metatheory
because their compose operators are structurally recursive, making
them easy to define in a proof assistant such as Agda. In contrast,
the coercions in normal form have compose operators that are not
structurally recursive, which makes it more difficult to define in
Agda, requiring what amounts to an explicit proof of termination.
%
On the other hand, coercions in normal form are easier to understand
than threesomes (with a strange labelled bottom type), and
supercoercions (with 10 different kinds).

%% There are two best studied ways to represent cast space-efficiently:
%% coercions
%% \citep{herman2010space}\citep{siek2009exploring}\citep{siek2012interpretations}
%% \citep{siek2015blame} and threesomes
%% \citep{siek2010threesomes}\citep{garcia2013calculating}. They have
%% different strength and weakness. \citet{garcia2013calculating} and
%% \citet{siek2015blame} claim that coercion is easy to understand, but
%% hard to implement, while threesome is easy to implement, but hard to
%% understand. \citet{garcia2013calculating} in addition claims that
%% coercion is more difficult to reason about formally because its
%% compose function is not structurally recurisive. For instance, the
%% developers of Grift \citep{kuhlenschmidt2018efficient} report that it
%% is tricky to convince their proof assistants that the compose function
%% terminates.
